Seed germination ecology of European mesic grasslands

### 2021-05-20 doc/manuscript.Rmd
new intro, discu, new figs 🌿

*Arrhenatheretalia*, plant regeneration, seed germination, seed morphology, species-rich meadows, hay meadows, mesic pastures

@@

European mesic grasslands are semi-natural open habitats that occupy moderately fertile and well-drained soils [@RN4698]. These plant communities developed by anthropogenic intervention in clearings of a formerly wooded landscape and are maintained by agricultural practices, contributing to the diversity and conservation value of European landscapes [@RN4996]. Traditional phytosociological classification of mesic grasslands put emphasis on the separation between mesic meadows used for hay making versus mesic pastures maintained by grazing, but a recent revision at the European level showed that the main driver of variation in species composition is the intensity rather than the type of management [@RN4893]. Indeed, changes in the intensity of management, including land abandonment and agricultural intensification, are threatening the maintenance of mesic grasslands in large parts of Europe [@RN4997]. For this reason the European Habitats Directive (92/43/EEC) has included in its lists of habitats of conservation interest those species-rich meadows that are traditionally managed by one or two annual cuts and light grazing.

One of the reasons for the high conservation interest of traditional mesic grasslands is their high species richness, and for this same reason they have been highlighted as a valuable source of natural seed materials to be used in ecological restoration, rewilding and urban greening [@RN4894]. However, a lack of knowledge about species germination traits has already been identified as a bottleneck [@RN2241] hampering the development of a competitive native seed industry [@RN4665] in Europe. Understanding seed germination is part of the practical scientific framework needed to tackle large-scale ecological restoration challenges [@RN3507] and to maintain ex situ collections of plant genetic resources [@RN3763]. When such a knowledge framework is available, ambitious regional schemes of seed-based landscape restoration can be designed [@RN4666].

In the case of mesic grasslands, a concerted investigation of their seed germination ecology is missing, despite their conservation and restoration importance. Most of the available information is scattered in single-species studies [@RN2236] or using a limited range of germination conditions [@RN4894]. In this article, we present for the first time a synthesis of germination ecology in European mesic grasslands. To do so, we combined a newly-generated dataset with germination records from ENSCOBASE, the seed germination database of the European Native Seed Conservation Network (http://enscobase.maich.gr/index.tml); and the *SylvanSeeds* database of seed germination records for nemoral biome species [@RN4968]. To the resulting dataset containing `r MSrecords` germination records from `r MScountries` European countries we applied Bayesian meta-analysis [@RN5000] to test the seed germination response to environmental cues including scarification, stratification, temperature, alternating temperature and light. In addition, we also analyzed seed morphology (mass and shape) and the differences between the main functional groups of mesic grasslands: grasses (Poaceae), legumes (Fabaceae) and the other families.

@@

In addition to this experimental data, we retrieved seed germination records from ENSCOBASE, the seed germination database of the European Native Seed Conservation Network (http://enscobase.maich.gr/index.tml); and the *SylvanSeeds* database of seed germination records for nemoral biome species [@RN4968]. The combined dataset, including our own experimental data and the records from ENSCOBASE and *SylvanSeeds*, contained `r MSrecords` germination records (i.e. germination proportions for a given seed lot of a species, recorded in a set of laboratory experimental conditions) from `r MScountries` European countries. There were `r MSspeciesN` species in the dataset. Overall, `r MSseedsN` seeds had been used in the experiments. The range of experimental germination temperatures (weighted average of the daily thermoperiod) which had been used in the experiments spans from `r MSminT` to `r MSmaxT` ºC, with `r MSaltN` records of constant temperatures (i.e. experiments that used the same temperature during all their duration) and `r MSaltY` of alternating temperatures (i.e. experiments where different temperatures were applied during the day and the night, in diurnal cycles). Seeds had been exposed to light during some part of the diurnal cycle in `r MSlightY` records or kept in total darkness in `r MSlightN` records. Experiments had been performed with unstratified seeds (i.e. not subjected to a previous dormancy-breaking incubation) in `r MSstratN` records and with stratified seeds (i.e. subjected to previous incubation in dormancy-breaking conditions, including treatments of wet incubation under cold, warm and combinations of cold and warm conditions) in `r MSstratY` records. Finally, there were `r MSgaY` records where GA~3~ had been applied, and `r MSscarY` records where seeds had been scarified.

@@

To test the effect of the germination treatments on seed germination proportions, we performed a meta-analysis of the germination dataset by fitting binomial generalized mixed models with Bayesian estimation (Markov Chain Monte Carlo generalized linear mixed models, MCMCglmms) using the R package *MCMCglmm* [@RN4755]. We fitted the analysis separately for the three major groups of mesic grassland plants: Poaceae, Fabaceae, and the minor families. To account for the effect of a shared phylogeny in species traits, models included as a random effect a reconstructed phylogenetic tree for the study species. To create the phylogeny we used the R package *V.PhyloMaker* [@RN4753] which contains an updated mega-tree of the seed plants based on Smith & Brown [-@RN4754]. We placed taxa absent from the mega-tree at the genus-level basal node. The phylogenetic tree is available in the data folder of the GitHub repository (see Data Availability Statement). Germination proportion was the response variable. Models had, as fixed effects, the experimental conditions (scarification, stratification, GA~3~, temperature, alternating temperature and light). Random effects included the phylogenetic tree, species identity, seed lot and source of the data. In all models, response variables were scaled so their contribution to the effect sizes could be compared. We used weakly informative priors in all models, with parameter-expanded priors for the random effects. Each model was run for 500,000 MCMC steps, with an initial burn-in phase of 50,000 and a thinning interval of 50 [@RN4756], resulting, on average, in 9,000 posterior distributions. From the resulting posterior distributions, we calculated mean parameter estimates and 95% Highest Posterior Density (HPD) and Credible Intervals (CI). We estimated the significance of model parameters by examining CIs, considering parameters with CIs overlapping with zero as non-significant. To estimate the phylogenetic signal of seed germination over all variables, we used Pagels’s lambda (λ) [@RN4757], estimated simultaneously with the models by calculating the mean of the posterior distribution and the 95% CI of λ as indicated by De Villemereuil et al. [-@RN4756]. When λ = 0, related taxa are no more similar than expected by chance, while when λ = 1, the trait is evolving following a constant variance random walk or Brownian motion model; intermediate values of λ indicate a phylogenetic correlation in trait evolution that does not fully follow a Brownian motion model [@RN4757]. Detailed results of the MCMCglmms are available in the results folder of the GitHub repository (see Data Availability Statement).

@@

The dispersal unit in the dominant grass family (Poaceae) is generally the floret (**Fig. 1a**), a composite structure made up of the caryopsis (a dry, indehiscent and monospermic fruit) and several surrounding bracts (i.e. modified leaves): the lemma, the palea and, in many species, two glumes. This structure is often accompanied by awns or hairs that aid in dispersal (**Fig. 1b**). The mass of these dispersal units in grassland grasses ranges from `r MSpoaMin` to `r MSpoaMax` mg, and they tend to be longer (`r MSpoaLength` mm on average) than they are wide (`r MSpoaWidth` mm). In the subordinate family of the legumes (Fabaceae), the dispersal unit of most species is the seed itself (**Fig. 1c**). Fabaceae generally are hard-seeded, i.e. they have a water-impermeable seed coat which needs to become permeable before germination can happen (i.e. physical seed dormancy). In some Fabaceae, dispersal units are more complex, including indehiscent monospermic fruits (e.g. *Onobrychis*, **Fig. 1d**) or indehiscent monospermic fruit fragments, i.e. loments (e.g. *Ornithopus*, **Fig. 1e**). In mesic grasslands, the dispersal units of the Fabaceae are generally heavier than those of the Poaceae, ranging from `r MSlegMin` to `r MSlegMax` mg; and they are also rounder (`r MSlegLength` mm long vs. `r MSlegWidth` mm wide). In many of the most frequent minor families, such as the Asteraceae, the Ranunculaceae, the Lamiaceae and the Dipsacaceae, the dispersal unit is the achene, i.e. a dry, indehiscent and monospermic fruit. In the case of the Asteraceae, these achenes tend to be cypselas, wearing a pappus that aids in dispersal by wind (**Fig. 1f**). In the Ranunculaceae, the achenes are usually flattened and sometimes beaked (**Fig. 1g**). The achenes of the Lamiaceae are hardened nutlets (**Fig. 1h**). In the genus *Knautia* (Dipsacaceae), the achene is covered by hairs (**Fig. 1i**). In the Cyperaceae, the dispersal unit is an achene surrounded by a perigynium which aids in dispersal by water (**Fig. 1j**). In the genus *Sanguisorba* (Rosaceae), the dispersal unit is the urn-shaped receptacle containing one to three achenes (**Fig. 1k**). The dispersal unit of the Apiaceae is the mericarp (**Fig. 1l**), an indehiscent monospermic fragment of the fruit. In some other minor families the dispersal unit is the seed itself, such as in the Caryophyllaceae (**Fig. 1m**) or the Plantaginaceae (**Fig. 1n**). The hemiparasitic species of *Rhinanthus* (Orobanchaceae) disperse by winged seeds (**Fig. 1o**). When comparing the morphology of all species (**Fig. 2**), the Poaceae appear as having relatively lighter and more elongated dispersal units, in contrast with the Fabaceae, which tend to have rounder and heavier seeds.

@@

* Mowing for haymaking thus is a predictable perturbation that occurs approximately at the same time every year, and which is expected to contribute to the high diversity of this habitats [@RN4887; @RN4888]. Hay meadow plants must establish, grow and reproduce in the time frame determined by this perturbation.

@@

>>>>

*Arrhenatheretalia*, *Arrhenatherion*, plant regeneration, seed germination, seed morphology, species-rich meadows, hay meadows, mesic pastures

@@

European mesic grasslands are semi-natural open habitats that occupy moderately fertile and well-drained soils [@RN4698]. These plant communities developed in clearings of a formerly temperate woodland landscape as a consequence of anthropogenic intervention [@RN5033]. At a time, they were novel plant communities with a composite flora made up of species from different origins: grasslands plants that existed in non-forested "islands" of the landscape, plus plants that evolved in parallel to human intervention by hybridisation or polyploidization [@RN5033]. During historical times the botanical composition of mesic grasslands evolved following changes in human activities and management practices [@RN5035]. They are thus the result of a process of habitat domestication that is characteristic of Europe's natural history [@RN5034]. Nowadays, mesic grasslands maintained by agricultural practices are a definitory aspect of European landscapes, both real [@RN4996] and imagined [@RN5029].

Traditional phytosociological classification of mesic grasslands puts emphasis on the separation between mesic meadows used for hay making versus mesic pastures maintained by grazing, but a recent revision at the European level showed that the main driver of variation in species composition is the intensity rather than the type of management [@RN4893]. Indeed, changes in the intensity of management, including land abandonment and agricultural intensification, are threatening the maintenance of mesic grasslands in large parts of Europe [@RN4997]. For this reason the European Habitats Directive (92/43/EEC) has included in its lists of habitats of conservation interest those species-rich meadows that are traditionally managed by one or two annual cuts and light grazing.

One of the reasons for the high conservation interest of traditional mesic grasslands is their high species richness, and for this same reason they have been highlighted as a valuable source of natural seed materials to be used in ecological restoration, rewilding and urban greening [@RN5030; @RN4894; @RN5031]. However, a lack of knowledge about species germination traits has already been identified as a bottleneck [@RN2241] hampering the development of a competitive native seed industry [@RN4665] in Europe. Understanding seed germination is part of the practical scientific framework needed to tackle large-scale ecological restoration challenges [@RN3507] and to maintain *ex situ* collections of plant genetic resources [@RN3763]. When such a knowledge framework is available, ambitious regional schemes of seed-based landscape restoration can be designed [@RN4666].

Mesic grasslands are dominated by mesophilous grasses of the Poaceae family, which make up most of the biomass and define the structure of the vegetation. The dispersal unit in the grasses is generally the floret (**Fig. 1a**), a composite structure made up of the caryopsis or grain (a dry, indehiscent and monospermic fruit) and its surrounding bracts (i.e. modified leaves): the lemma and the palea. Removal of the lemma and palea, as well as puncturing the pericarp, can aid in seed germination [@RN1248]. This structure is often accompanied by awns or hairs (**Fig. 1b**) that are assumed to aid dispersal, although it is not always the case that they do [@RN1290]. Grass seeds have been reported as having physiological seed dormancy [@RN5018; @RN3214; @RN5024] in various degrees, but germinability is usually high even without treating the seeds with cold stratification [@RN1205; @RN1260; @RN1271; @RN1229; @RN372; @RN789; @RN680; @RN1290; @RN398; @RN223; @RN334; @RN1281; @RN5024; @RN3273]. Freshly harvested seeds are comparatively more dormant, but dormancy tends to disappear quickly in dry storage [@RN5024; @RN5018]. Germination has been reported to occur at temperatures ranging from 5 to 30 ºC [@RN3273; @RN1229; @RN1261; @RN1230; @RN1292; @RN5024]. Most Poaceae species have also been reported to germinate better in light than in darkness [@RN1271; @RN1248; @RN1230; @RN1221; @RN1172; @RN1292; @RN1280; @RN5024] and even to be inhibited to germinate by darkness [@RN1229]. Nonetheless, species of *Bromus* have been reported as germinating better in darkness [@RN1172] and *Cynosurus cristatus* as being indifferent to light/darkness [@RN1292; @RN1280]. In *Poa trivialis*, germination was promoted by light but not by alternating temperatures [@RN1205], although alternating temperatures did encourage some germination in darkness [@RN1229]. Wild grass ecotypes usually have a germination response to alternating temperatures [@RN1248; @RN1261; @RN1230; @RN1221; @RN1290; @RN1172; @RN1292; @RN1280], although this response is missing in some wild ecotypes and in the domesticated cultivars [@RN1261], and some species such as *Lolium perenne* have been reported as insensitive to temperature alternation [@RN1280; @RN1380].

Next in abundance to the grasses are the legumes of the Fabaceae family. Legumes contribute to the nutritional value of grassland fodder, as thanks to their N-fixating capabilities they have high N contents [@RN5028; @RN5040]. The dispersal unit of most legumes is the seed itself (**Fig. 1c**) but in some species dispersal units are more complex, including indehiscent monospermic fruits (e.g. *Onobrychis*, **Fig. 1d**) or indehiscent monospermic fruit fragments, i.e. loments (e.g. *Ornithopus*, **Fig. 1e**). Legumes are generally hard-seeded, i.e. they have a water-impermeable seed coat which needs to become permeable before germination can happen (i.e. physical seed dormancy) [@RN3214; @RN3273; @RN1215; @RN5016; @RN3073]. In *Medicago*, seeds that have not reached full maturity can germinate and are not yet impermeable, but the completion of maturation imposed coat impermeability, and thereafter the seed coat must be scarified to allow water imbibition and germination [@RN652]. Buried legume seeds are able to track the seasons, and in some species germination seems to be promoted by cold stratification and alternating temperatures [@RN780], some of these have been described as having combinational dormancy (i.e. physical + physiological) [@RN5037]. However, seeds have also been reported to germinate without any previous treatment [@RN881; @RN632; @RN460; @RN372], and to lose dormancy during storage [@RN5037]. As in the grasses, legume seeds have been reported to germinate in high numbers across a range of temperatures from 5 to 25 ºC [@RN652; @RN3273]. Legume seeds have been described as not responsive to light and capable of germinating in darkness [@RN1339; @RN3273].

Grasses and legumes are accompanied by a diversity of other families which, even if present in lower abundances, contribute to the aesthetic value of the meadows [@RN5038; @RN5039]. They also add nutritional quality, being richer than grasses and legumes in specific elements [@RN5028; @RN5040]. In many of the most frequent minor families (e.g. in the Asteraceae, Ranunculaceae, Polygonaceae, Lamiaceae, Dipsacaceae, Cyperaceae) the dispersal unit is the achene: dry, indehiscent and monospermic fruits. The morphology of these achenes is varied: cypselas with a hairy pappus in *Centaurea* (Asteraceae, **Fig. 1f**); beaked in *Ranunculus* (Ranunculaceae, **Fig. 1g**); hardened nutlets in *Prunella* (Lamiaceae, **Fig. 1h**); hairy in *Knautia* (Dipsacaceae, **Fig. 1i**); surrounded by a perigynium which aids in dispersal by water in *Carex* (Cyperaceae, **Fig. 1j**). In the genus *Sanguisorba* (Rosaceae), the dispersal unit is the urn-shaped receptacle containing one to three achenes (**Fig. 1k**). In the Apiaceae, it is the mericarp (**Fig. 1l**), an indehiscent monospermic fragment of the fruit. In some other minor families the dispersal unit is the seed itself, such as in the Caryophyllaceae (**Fig. 1m**), the Plantaginaceae (**Fig. 1n**), the Juncaceae or the hemiparasitic species of *Rhinanthus* (Orobanchaceae) (**Fig. 1o**). Our understanding of the germination ecology of these seeds is as varied as their morphology. In the Asteraceae, high germination without pretreatments has been reported in *Taraxacum officinale* [@RN1276; @RN194; RN120; @RN963; @RN1428] and *Hypochaeris radicata* [@RN194; @RN1403] and *Achillea millefolium* [@RN1403]. *Taraxacum officinale* germinated between 5 and 30 ºC [@RN120; RN1428; @RN1276], and had higher germination in light [@RN1015; @RN963; @RN1172] and in alternating temperatures [@RN1428]. In *Stachys officinalis* (Lamiaceae), seeds need either cold stratification, light or alternating temperatures to germinate [@RN416; @RN128]. Underdevelopped embryos that need to grow inside the seed before germination (i.e. morphological dormancy) are widespread in the Ranunculaceae and the Apiaceae [@RN3214; @RN1112]. *Ranunculus repens* (Ranunculaceae) germinated between 10 and 25 ºC, but the germination percentages were low [@RN931]; the same species has been reported to respond to alternating temperatures, which can promote its germination even in darkness [@RN3180]. In the Polygonaceae, *Rumex acetosa* can germinate inmediately after dispersal and between 7 and 27 ºC, while the congeneric *Rumex acetosella* does not, this differences is due to the former being able to germinate at constant temperatures in the darkness [@RN814; @RN3273], while the latter has an absolute requirement for light [@RN814]. In *Heracleum sphondylium* (Apiaceae), growth of the embryo only occurs below 10 ºC, in moist conditions [@RN1112]. *Sanguisorba minor* (Rosaceae) increased its germination after abrasion of the seeds with bleach [@RN194; @RN226], although germination without pre-treatment has also been reported [@RN298; @RN226]. Seeds of the hemiparasitic species *Rhinanthus angustifolius* and *Rhinanthus minor* (Orobanchaceae) required relatively long periods of cold stratification to germinate [@RN4014], and were able to germinate in the dark [@RN2236].

Although a wealth of studies has accumulated, a synthesis of the seed germination ecology of mesic grasslands is missing, despite their conservation, restoration and cultural significance. In this article, we present for the first time a synthesis of germination ecology in European mesic grasslands. To do so, we combined a newly-generated dataset with germination records from ENSCOBASE [@RN5036], the seed germination database of the European Native Seed Conservation Network (http://enscobase.maich.gr/index.tml); and the *SylvanSeeds* database of seed germination records for nemoral biome species [@RN4968]. To the resulting dataset containing `r MSrecords` germination records from `r MScountries` European countries we applied Bayesian meta-analysis [@RN5000] to test the seed germination response to environmental cues including scarification, stratification, temperature, alternating temperature and light. In addition, we also analyzed seed morphology (mass and shape) and the differences between the main functional groups of mesic grasslands: grasses (Poaceae), legumes (Fabaceae) and the other families.

@@

In addition to this experimental data, we retrieved seed germination records from ENSCOBASE [@RN5036], the seed germination database of the European Native Seed Conservation Network (http://enscobase.maich.gr/index.tml); and the *SylvanSeeds* database of seed germination records for nemoral biome species [@RN4968]. The combined dataset, including our own experimental data and the records from ENSCOBASE and *SylvanSeeds*, contained `r MSrecords` germination records (i.e. germination proportions for a given seed lot of a species, recorded in a set of laboratory experimental conditions) from `r MScountries` European countries. There were `r MSspeciesN` species in the dataset. Overall, `r MSseedsN` seeds had been used in the experiments. The range of experimental germination temperatures (weighted average of the daily thermoperiod) which had been used in the experiments spans from `r MSminT` to `r MSmaxT` ºC, with `r MSaltN` records of constant temperatures (i.e. experiments that used the same temperature during all their duration) and `r MSaltY` of alternating temperatures (i.e. experiments where different temperatures were applied during the day and the night, in diurnal cycles). Seeds had been exposed to light during some part of the diurnal cycle in `r MSlightY` records or kept in total darkness in `r MSlightN` records. Experiments had been performed with unstratified seeds (i.e. not subjected to a previous dormancy-breaking incubation) in `r MSstratN` records and with stratified seeds (i.e. subjected to previous incubation in dormancy-breaking conditions, including treatments of wet incubation under cold, warm and combinations of cold and warm conditions) in `r MSstratY` records. Finally, there were `r MSgaY` records where GA~3~ had been applied, and `r MSscarY` records where seeds had been scarified.

@@

To test the effect of the germination treatments on seed germination proportions, we performed a meta-analysis of the germination dataset by fitting binomial generalized mixed models with Bayesian estimation (Markov Chain Monte Carlo generalized linear mixed models, MCMCglmms) [@RN5036; @RN4965] using the R package *MCMCglmm* [@RN4755]. We fitted the analysis separately for the three major groups of mesic grassland plants: Poaceae, Fabaceae, and the minor families. To account for the effect of a shared phylogeny in species traits, models included as a random effect a reconstructed phylogenetic tree for the study species. To create the phylogeny we used the R package *V.PhyloMaker* [@RN4753] which contains an updated mega-tree of the seed plants based on Smith & Brown [-@RN4754]. We placed taxa absent from the mega-tree at the genus-level basal node. The phylogenetic tree is available in the data folder of the GitHub repository (see Data Availability Statement). Germination proportion was the response variable. Models had, as fixed effects, the experimental conditions (scarification, stratification, GA~3~, temperature, alternating temperature and light). Random effects included the phylogenetic tree, species identity, seed lot and source of the data. In all models, response variables were scaled so their contribution to the effect sizes could be compared. We used weakly informative priors in all models, with parameter-expanded priors for the random effects. Each model was run for 500,000 MCMC steps, with an initial burn-in phase of 50,000 and a thinning interval of 50 [@RN4756], resulting, on average, in 9,000 posterior distributions. From the resulting posterior distributions, we calculated mean parameter estimates and 95% Highest Posterior Density (HPD) and Credible Intervals (CI). We estimated the significance of model parameters by examining CIs, considering parameters with CIs overlapping with zero as non-significant. To estimate the phylogenetic signal of seed germination over all variables, we used Pagels’s lambda (λ) [@RN4757], estimated simultaneously with the models by calculating the mean of the posterior distribution and the 95% CI of λ as indicated by De Villemereuil et al. [-@RN4756]. When λ = 0, related taxa are no more similar than expected by chance, while when λ = 1, the trait is evolving following a constant variance random walk or Brownian motion model; intermediate values of λ indicate a phylogenetic correlation in trait evolution that does not fully follow a Brownian motion model [@RN4757]. Detailed results of the MCMCglmms are available in the results folder of the GitHub repository (see Data Availability Statement).

@@

The mass of these dispersal units in grassland grasses ranges from `r MSpoaMin` to `r MSpoaMax` mg, and they tend to be longer (`r MSpoaLength` mm on average) than they are wide (`r MSpoaWidth` mm).

In mesic grasslands, the dispersal units of the Fabaceae are generally heavier than those of the Poaceae, ranging from `r MSlegMin` to `r MSlegMax` mg; and they are also rounder (`r MSlegLength` mm long vs. `r MSlegWidth` mm wide).

When comparing the morphology of all species (**Fig. 2**), the Poaceae appear as having relatively lighter and more elongated dispersal units, in contrast with the Fabaceae, which tend to have rounder and heavier seeds.

@@

Our meta-analysis of germination records showed that, overall, the environmental factors with the strongest effect to control seed germination in mesic grassland species were alternating temperatures, light and scarification. Alternating temperatures and light are usually considered to be micro-environmental cues that indicate, at a fine scale, the existence of safe sites for regeneration [@RN4719]. The diurnal alternation of temperatures decreases with burial depth in the soil, and the alternation is also reduced by vegetation cover [@RN1380; @RN4910; @RN4952]. Thus, a positive germination response to alternating temperatures could detect the depth of seed burial, and prevent the germination of seeds that are buried too deep for seedlings to survive before emerging from the ground [@RN4898]. Perhaps more importantly in this habitat, alternating temperatures could also indicate that the vegetation cover has been diminished by either sowing or grazing, and therefore mark an appropriate time for seedling regeneration, when the competition by the established plants is lessened. Similar functions can be ascribed to the germination response to light, as physiologically-active light in significant amounts only travels the uppermost milliliters of soil [@RN4909] and the quality of light will be affected by vegetation cover [@RN604]. Therefore, the germination response to alternating temperatures and light indicate conditions that are found immediately after mowing for haymaking, a predictable perturbation that occurs approximately at the same time every year, and which forces meadow plants to establish, grow and reproduce in the time frame determined by this perturbation [@RN4887; @RN4888]. In addition, this traits would also promote the formation of a persistent soil seed bank [@RN4718; @RN4929; @RN1292]. On the other hand, it is worth mentioning that the Poaceae, which show the germination response to alternating temperatures and light, also have the more elongated seed shape, a trait that is thought to reduce the capacity of seeds to enter the soil seed bank in temperate grasslands [RN5014; @RN5013].

The strong effect of scarification was specifically related to the hard-seeded Fabaceae. In this family, physical dormancy as a result of an unpermeable seed coat works as a mechanism to detect seasonal cycles of temperature and humidity [@RN780]. It has been also proposed that hardseedeness can avoid seed predation, by preventing the emission of olfactory cues that are elicited by seed imbibition and that can be perceived by seed predators [@RN2568]. In the case of this study, Fabaceae seeds are the heaviest, and they clearly follow a separate regeneration strategy from that of the other families. While other groups rely on alternating temperatures and light as gap-detecting mechanisms, Fabaceae seeds do not respond to these factors, and instead they are mainly regulated by scarification, and they also respond to cooler temperatures than the rest of species. This would indicate that Fabaceae seeds tend to germinate when temperatures are cooler, and thus before or after the summer hay-making season. Their larger size, and the related larger reserves, could allow Fabaceae seedlings to emerge at times when the competition from the established vegetation would be higher.

The germination response to average temperatures and stratification is understood to detect cues related to seasonal cycling and macroclimatic variation [@RN3063, @RN5036]. Apart from the aforementioned response to cool temperatures in the Fabaceae, the rest of the families did not show a response to average temperatures, further highlighting their reliance on micro-habitat cueing. Moreover, the dominant family of the Poaceae showed a negative response to cold stratification. The minor families, however, showed a divergence from the Poaceae strategy in their positive response to stratification. These subordinate species appear to require a period of cold stratification, indicating that overwintering has occured, before they can germinate [@RN3214]. These minor families are also the only group that shows a positive response to GA~3~, a phytohormone which can work as a substitute of cold stratification to overcome physiological dormancy [@RN3368].

The positive phylogenetic signal found in the models highlights the phylogenetic clustering of regeneration strategies that we have described so far: (1) the dominant family of the Poaceae, showing a lack of responses to seasonal cues (average temperature and cold stratification), and relying instead on micro-niche cues (alternating temperatures, light) that can be associated to the yearly perturbation of mowing, (2) the second most-dominant family, the Fabaceae, which do not respond to the micro-niche cues, and regulate their germination timing through scarification and cooler germination temperatures, possibly because their larger size and reserves allow their seeds to decouple their emergence timing from the mowing perturbance, and finally (3) the minor families that respond to micro-cues in a similar way than the Poaceae, but which are differentieted from them by showing a positive response to cold stratification and G~3~, indicating that they rely on physiological seed dormancy.

* Less dormancy in grassland vs. arable populations [@RN1229]. Herbaceous plants adapted to open habitats needed less cold stratification than congeneric species adapted to forest understories [@RN3570].

* Better germination without treatments [@RN4894].

* In Poland soil bank dominated by forbs of arable and weedy species, low representation of grasses and legumes, confirms the typical Arrhenatherion species do not form a seed bank [@RN5032]. Soil bank is transient in hay meadows and related grasslands [@RN5026; @RN5041; @RN5042]. Regeneration limited by dispersal limitation [@RN3073].

* Mowing promotes germination and emergence [@RN3073]. Mowing promotes autumn germination [@RN5025].

@@

* Conclusions/applications

* Discuss dormancy in freshly collected grasses, limitations of stored seed data [@RN5018]
